Efficient Compilation of High Level Python Numerical Programs with Pythran

### 2014-09-03 doc/papers/cgo2014/paper.tex
update graphes

% For performance comparison between different Pythran version
To compare the different optimization schemes, we run the numpy-benchmarks test
suite multiple times on the \textit{rosen} benchmark. The algorithm is the one
on Listing~\ref{lst:rosen} and input is a raw array of 1,000,000 of double elements.
Figures~\ref{fig:rosenamd48} and \ref{fig:rosenintel8} illustrate the results.
The speedup difference between the two architectures can be explain by AVX support on
the Intel8 architecture while AMD48 supports up to SSE4 only.
We can also notice importance of forward substitution on this benchmarks as
we avoid extra loop, 2 temporary array assignment and avoid some load/store for
SIMD instructions.

@@

\subsection{Comparison with others Python compiler}

Figure~\ref{fig:arc_distance} is a typical example where we can make intensive
use of the techniques described in this paper. It measures performances on the
arc\_distance benchmark. It is a kernel with intensive points to points computation
and trigonometrics use. As we can see, both vectorization
and parallelization have an impact on execution time. We can notice that SIMD
contribution is not as good on AMD48 as it is on Intel8, because it does not
support AVX. This program also benefits from OpenMP. Input is 10000 elements 
double raw array.

%FIXME: mehdi propose de n'utiliser qu'un backend pour centrer le propos sur Pythran/Numpy.
% ça allègera aussi les graphes...
% On change juste le jeu de donnée plutot que de supprimer une archi

\begin{figure}[t]
\centering
\includegraphics[width=0.5\textwidth]{arc_distance_amd48.png}
\caption{Execution time of the arc distance kernel on AMD48.}
\label{fig:arc_distance}
\end{figure}

\begin{figure}[t]
\centering
\includegraphics[width=0.5\textwidth]{arc_distance_intel8.png}
\caption{Execution time of the arc distance kernel on Intel8.}
\label{fig:arc_distance_intel8}
\end{figure}

Figure~\ref{fig:diffusion} show how Clang generated code is more performant
when handling \lstinline{gsliced_ndarray}.
This structure is a non aligned view on an array which prevent SIMD optimization
on the second dimension even if loop fusion and temporary elimination is
performed.  We can notice that Parakeet performs really well on
extensive gslice use.
%TODO IL faut refaire les mesures en changeant la limite OpenMP. Parce qu'on a
%pas du tout de parallelisme la. On note par contre l"overhead libgomp...
%SG: +1

\begin{figure}[t]
\centering
\includegraphics[width=0.5\textwidth]{diffusion_amd48.png}
\caption{Execution time of the diffusion kernel on AMD48.}
\label{fig:diffusion}
\end{figure}

\begin{figure}[t]
\centering
\includegraphics[width=0.5\textwidth]{diffusion_intel8.png}
\caption{Execution time of the diffusion kernel on Intel8.}
\label{fig:diffusion_intel8}
\end{figure}

Finally, Figure~\ref{fig:l2norme} shows the L2norm algorithm introduced in
\S~\ref{sec:stackoverflow}. Comparison with other compilers illustrates how the
approach taken by Pythran is relevant for these kind of functions.
%TODO IL faut refaire les mesures en changeant la limite OpenMP. Parce qu'on a
%pas du tout de parallelisme la. On note par contre l"overhead libgomp...
%SG: +1
\begin{figure}[t]
\centering
\includegraphics[width=0.5\textwidth]{l2norm_amd48.png}
\caption{Execution time of the l2norm kernel on AMD48.}
\label{fig:l2norme}
\end{figure}

>>>>

% For performance comparison between different Pythran version
To compare the different optimization schemes, we run the numpy-benchmarks test
suite multiple times on the \textit{rosen} benchmark. The algorithm is the one
on Listing~\ref{lst:rosen} and input is a raw array of 1,000,000 double elements.
Figures~\ref{fig:rosenamd48} and \ref{fig:rosenintel8} illustrate the results.
The speedup difference between the two architectures can be explain by AVX support on
the Intel8 architecture while AMD48 supports up to SSE4 only.
We can also notice importance of forward substitution on this benchmarks as
we avoid extra loop, 2 temporary array assignment and avoid some load/store for
SIMD instructions.

@@

\subsection{Comparison with others Python compiler}
\todo{a better title have to be found}

Figure~\ref{fig:arc_distance} and \ref{fig:arc_distance_intel8} are typicals
example where we can make intensive
use of the techniques described in this paper. It measures performances on the
arc\_distance benchmark. It is a kernel with intensive points to points computation
and trigonometrics use. As we can see, both vectorization
and parallelization have an impact on execution time. We can notice that SIMD
contribution is not as good on AMD48 as it is on Intel8, because it does not
support AVX. This program also benefits from OpenMP and is scalable up to 20 cores.
The input is a raw array of 10,000,000 double elements.

\begin{figure}[t]
\centering
\includegraphics[width=0.5\textwidth]{arc_distance_amd48.png}
\caption{Speed up for the arc distance kernel on AMD48.}
\label{fig:arc_distance}
\end{figure}

\begin{figure}[t]
\centering
\includegraphics[width=0.5\textwidth]{arc_distance_intel8.png}
\caption{Speed up for the arc distance kernel on Intel48.}
\label{fig:arc_distance_intel8}
\end{figure}

Finally, Figure~\ref{fig:l2norme} and \ref{fig:l2norm_intel8} shows the L2norm algorithm introduced in
\S~\ref{sec:stackoverflow}. Comparison with other compilers illustrates how the
approach taken by Pythran is relevant for these kind of functions. Just like for
the arc\_distance benchmarks, we can notice use of AVX when it is available.

\begin{figure}[t]
\centering
\includegraphics[width=0.5\textwidth]{l2norm_amd48.png}
\caption{Execution time of the l2norm kernel on AMD48.}
\label{fig:l2norme}
\end{figure}
